A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-07-11 paper/prab.tex
starting to fix verbatim sections

\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{todonotes}
\usepackage{siunitx}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\usepackage{booktabs}
%\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{url}
%\usepackage{footmisc}
%\usepackage{multicol}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,snakes,backgrounds}
\usetikzlibrary{mindmap,trees}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{plotmarks}
\usepackage{listings}
\newcounter{lstmain}
\setcounter{lstmain}{1}

@@

To allow arbitrary constraints and objective expressions, such as
%
\vspace{0.2cm}
{\footnotesize \begin{verbatim}
  name: OBJECTIVE,
        EXPR="5 * average(42.0, "measurement.dat") + ENERGY";
\end{verbatim}}
\vspace{0.2cm}
%
\noindent
  we implemented an expression parser using Boost
%  Spirit\footnote{\url{http://boost-spirit.com/}}.
In addition to the parser, we need an evaluator able to evaluate an expression,
  given a parse tree and variable assignments to an actual value.
Expressions arising in multi-objective optimization problems usually evaluate
  to booleans or floating point values.
The parse tree, also denoted abstract syntax tree (AST), is constructed
  recursively while an expression is parsed.
Upon evaluation, all unknown variables are replaced with values, either
  obtained from simulation results or provided by other subtrees in the AST.
In this stage, the AST can be evaluated bottom-up and the desired result is
  returned after processing the root of the tree.

@@

\vspace{0.2cm}
{\footnotesize \begin{verbatim}
  d1: DVAR, VARIABLE="x1", LOWERBOUND="-1.0", UPPERBOUND="1.0";
  d2: DVAR, VARIABLE="x2", LOWERBOUND="-1.0", UPPERBOUND="1.0";
  d3: DVAR, VARIABLE="x3", LOWERBOUND="-1.0", UPPERBOUND="1.0";

@@

  objs:    OBJECTIVES = (obj1, obj2);
  dvars:   DVARS = (d1, d2, d3);
  constrs: CONSTRAINTS = ();
  opt: OPTIMIZE, OBJECTIVES=objs, DVARS=dvars, CONSTRAINTS=constrs;
\end{verbatim}}
\vspace{0.2cm}

>>>>

\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{todonotes}
\usepackage{siunitx}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\usepackage{fancyvrb}
\usepackage{booktabs}
%\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{url}
%\usepackage{footmisc}
%\usepackage{multicol}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,snakes,backgrounds}
\usetikzlibrary{mindmap,trees}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{plotmarks}
\usepackage{listings}
\newcounter{lstmain}
\setcounter{lstmain}{1}

@@

To allow arbitrary constraints and objective expressions, such as
%
\vspace{0.2cm}
\begin{Verbatim}[fontsize=\scriptsize]
  name: OBJECTIVE,
        EXPR="5 * average(42.0, "measurement.dat") + ENERGY";
\end{Verbatim}
\vspace{0.2cm}
%
\noindent
  we implemented an expression parser using Boost
%  Spirit\footnote{\url{http://boost-spirit.com/}}.
In addition to the parser, we need an evaluator able to evaluate an expression,
  given a parse tree and variable assignments to an actual value.
Expressions arising in multi-objective optimization problems usually evaluate
  to booleans or floating point values.
The parse tree, also denoted abstract syntax tree (AST), is constructed
  recursively while an expression is parsed.
Upon evaluation, all unknown variables are replaced with values, either
  obtained from simulation results or provided by other subtrees in the AST.
In this stage, the AST can be evaluated bottom-up and the desired result is
  returned after processing the root of the tree.

@@

\vspace{0.2cm}
\begin{flushleft}
\noindent
\begin{Verbatim}[fontsize=\scriptsize]
d1: DVAR, VARIABLE="x1", LOWERBOUND="-1.0", UPPERBOUND="1.0";
d2: DVAR, VARIABLE="x2", LOWERBOUND="-1.0", UPPERBOUND="1.0";
d3: DVAR, VARIABLE="x3", LOWERBOUND="-1.0", UPPERBOUND="1.0";

@@

objs:    OBJECTIVES = (obj1, obj2);
dvars:   DVARS = (d1, d2, d3);
constrs: CONSTRAINTS = ();
opt: OPTIMIZE, OBJECTIVES=objs, DVARS=dvars, CONSTRAINTS=constrs;
\end{Verbatim}
\end{flushleft}

\vspace{0.2cm}

### 2018-07-11 paper/theor-experiment.tex
starting to fix verbatim sections

\begin{figure}
  \centering
  \input{figures/ferrario_illustrated}
  \caption{Illustration of the Ferrario matching criteria: beam emittance
  attains a maximum and rms beamsize a minimum at the entrance to the first
  accelerating traveling wave structure.}
  \label{fig:fer_match}
\end{figure}

@@

In order to compute the peaks, we employed an additional Python script.
This script was called in the \textsc{OPAL}~input file, after the simulation
  finished using the \texttt{SYSTEM} functionality.
Once the peaks (in a given range) were located, the two objectives
  (\ref{eq:swissfel:p1}) and (\ref{eq:swissfel:p2}) were computed and their
  values written into corresponding files.
The custom \texttt{fromFile} functor allows us to access the values stored in
  the peak finder Python script result files

\vspace{0.2cm}
{\footnotesize \begin{verbatim}
    rmsx:  OBJECTIVE, EXPR="statVariableAt("rms_x-err.dat", "var")";
    emitx: OBJECTIVE, EXPR="statVariableAt("emit_x-err.dat", "var")";
    match: OBJECTIVE, EXPR="fabs(statVariableAt("emit_x-peak.dat", "var") -
                                 statVariableAt("rms_x-peak.dat", "var"))";
\end{verbatim}}
\vspace{0.2cm}

\noindent
The design variables and the assembly of the multi-objective optimization problem
  can be included in the \textsc{OPAL}~input file as shown below:

\vspace{0.2cm}
{\footnotesize \begin{verbatim}
    d1:  DVAR, VARIABLE="SIGX", LOWERBOUND="0.00025", UPPERBOUND="0.00029";
    d2:  DVAR, VARIABLE="FIND1_MSOL10_i", LOWERBOUND="110", UPPERBOUND="120";
    d3:  DVAR, VARIABLE="D_LAG_RGUN", LOWERBOUND="-0.1", UPPERBOUND="0.1";
    d4:  DVAR, VARIABLE="D_SOLPOS", LOWERBOUND="-0.05", UPPERBOUND="0.05";

    objs:    OBJECTIVES = (rmsx, emitx);
    dvars:   DVARS = (d1, d2, d3, d4);
    constrs: CONSTRAINTS = ();
    opt:     OPTIMIZE, OBJECTIVES=objs, DVARS=dvars,
             CONSTRAINTS=constrs;
\end{verbatim}}
\vspace{0.2cm}

@@

Each of the $40$ points on the Pareto front, shown in
  Figure~\ref{fig:pareto_front_21}, represents an optimal solution, where
  emittance and beamsize values are compromised to achieve the best agreement
  with the Ferrario matching point.
We selected individual $3$ based on a comparison of the emittance and beamsize
  characteristics of all solutions and by retaining the feasibility of the
  beam line optics parameters.
The design variables, emittance and beamsize of the selected solution are
  shown in Table~\ref{tbl:des_vars} and Figure~\ref{fig:rmsemit}.
With the multi-objective optimization framework we attain the same working
  point as reported in~\cite{pedr:10}.

@@

\begin{figure}%[h!]
  \centering
  \includegraphics[width=0.8\linewidth]{Report/iff_plot_emrms}
  \caption{Beamsize and emittance of individual $3$.}
  \label{fig:rmsemit}
\end{figure}

@@

For validation purposes we compared the results of the envelope-tracker using
  the analytical space charge model with the \textsc{OPAL} 3D macro particle
  tracker.
The benchmark was run on the first 12.5 meters of the \textsc{SwissFEL} 250
  MeV injector.
The results for both rms beamsize and emittance are shown in
  Figure~\ref{fig:et-vs-tt}.
A good agreement between the two codes can be observed.
The difference of the larger emittance along the solenoids in case of 3D
  tracker that is not seen by the envelope-tracker is due to the different
  definition of the particle momenta (canonical vs.~mechanical).
Both trackers agree within acceptable limits \cite{chao:99}.

>>>>

\begin{figure}
  \centering
  \input{figures/ferrario_illustrated}
  \caption{Illustration of the Ferrario matching criteria: beam emittance
  attains a maximum and rms beam size a minimum at the entrance to the first
  accelerating traveling wave structure.}
  \label{fig:fer_match}
\end{figure}

@@

In order to compute the peaks, we employed an additional Python script.
This script was called in the \textsc{OPAL}~input file, after the simulation
  finished using the \texttt{SYSTEM} functionality.
Once the peaks (in a given range) were located, the two objectives
  (\ref{eq:swissfel:p1}) and (\ref{eq:swissfel:p2}) were computed and their
  values written into corresponding files.
The custom \texttt{fromFile} functor allows us to access the values stored in
  the peak finder Python script result files
%
%\vspace{0.2cm}
\begin{flushleft}
\noindent
\begin{Verbatim}[fontsize=\scriptsize]
rmsx:  OBJECTIVE, EXPR="statVariableAt("rms_x-err.dat", "var")";
emitx: OBJECTIVE, EXPR="statVariableAt("emit_x-err.dat", "var")";
match: OBJECTIVE, EXPR="fabs(statVariableAt("emit_x-peak.dat", "var") -
statVariableAt("rms_x-peak.dat", "var"))";
\end{Verbatim}
\end{flushleft}
\vspace{0.2cm}

\noindent
The design variables and the assembly of the multi-objective optimization problem
  can be included in the \textsc{OPAL}~input file as shown below:
%\vspace{0.2cm}
\begin{flushleft}
\noindent
\begin{Verbatim}[fontsize=\scriptsize]
d1: DVAR, VARIABLE="SIGX", LOWERBOUND="0.00025", UPPERBOUND="0.00029";
d2: DVAR, VARIABLE="FIND1_MSOL10_i", LOWERBOUND="110", UPPERBOUND="120";
d3: DVAR, VARIABLE="D_LAG_RGUN", LOWERBOUND="-0.1", UPPERBOUND="0.1";
d4: DVAR, VARIABLE="D_SOLPOS", LOWERBOUND="-0.05", UPPERBOUND="0.05";

objs:    OBJECTIVES = (rmsx, emitx);
dvars:   DVARS = (d1, d2, d3, d4);
constrs: CONSTRAINTS = ();
opt:     OPTIMIZE, OBJECTIVES=objs, DVARS=dvars,
CONSTRAINTS=constrs;
\end{Verbatim}
\end{flushleft}
\vspace{0.2cm}

@@

Each of the $40$ points on the Pareto front, shown in
  Figure~\ref{fig:pareto_front_21}, represents an optimal solution, where
  emittance and beam size values are compromised to achieve the best agreement
  with the Ferrario matching point.
We selected individual $3$ based on a comparison of the emittance and beam size
  characteristics of all solutions and by retaining the feasibility of the
  beam line optics parameters.
The design variables, emittance and beam size of the selected solution are
  shown in Table~\ref{tbl:des_vars} and Figure~\ref{fig:rmsemit}.
With the multi-objective optimization framework we attain the same working
  point as reported in~\cite{pedr:10}.

@@

\begin{figure}%[h!]
  \centering
  \includegraphics[width=0.8\linewidth]{Report/iff_plot_emrms}
  \caption{Beam size and emittance of individual $3$.}
  \label{fig:rmsemit}
\end{figure}

@@

For validation purposes we compared the results of the envelope-tracker using
  the analytical space charge model with the \textsc{OPAL} 3D macro particle
  tracker.
The benchmark was run on the first 12.5 meters of the \textsc{SwissFEL} 250
  MeV injector.
The results for both rms beam size and emittance are shown in
  Figure~\ref{fig:et-vs-tt}.
A good agreement between the two codes can be observed.
The difference of the larger emittance along the solenoids in case of 3D
  tracker that is not seen by the envelope-tracker is due to the different
  definition of the particle momenta (canonical vs.~mechanical).
Both trackers agree within acceptable limits \cite{chao:99}.
